Designing and Modeling a Browser-Based Distributed Evolutionary Computation System

### 2015-05-09 GECCO/wk0705.tex
Including base results

% Rewrite this - JJ
In order to test the volunteer computing environment, a presentation
describing a low cost volunteer computing environment was
created. This presentation was actually delivered in several
conferences (such as IDC 2014, PPSN 2014 and FOSDEM 2015). These
results have been presented in \cite{arxiv:modeling}. In this paper,
although it uses exactly the same evolutionary algorithm client code,
the {\em presentation} has changed: it uses a single page with several
panels showing the progress of the algorithm and how many users have
been contributing to the experiment. To get users, the URL
(\url{http://nodio-jmerelo.rhcloud.com} and
\url{http://nodeo-i.herokuapp.com}) were presented, with a short
message, in Twitter and other social networks, introducing it as a
``demo for evolutionary algorithms'', sometimes in Spanish, some in
English.

@@

\section{Results}
\label{sec:res}

The experiments done with the first versions of this framework have
already been published elsewhere \cite{arxiv:modeling}. We proved that
up to 20 nodes could participate in a single run if it was
sufficiently announced; every experiment was individually announced so
more or less a certain amount of nodes were guaranteed. However, it
did not run correctly in mobile phones, design of the presentation was
not adaptive and thus missed a good amount of possible clients. In
this case the design, which has used the Kube CSS framework, was fully
adaptive and indeed proved to work with mobile phones browsers for
iPhones and Android phones. 
%
\begin{figure}[h!tb]
  \includegraphics[width=0.95\linewidth]{img/openshift-intervals.png}
  \caption{Histogram of the intervals between two requests by the same
  IP for all experiments in the OpenShift server.}  \label{fig:intervals}
\end{figure}
%
\begin{figure*}[h!tb]
  \includegraphics[width=0.90\linewidth]{img/openshift-intervals-IP.png}
  \caption{Boxplot of the intervals between two requests by the same
  IP, by IP, ordered by descending median.}  \label{fig:intervalIP}
\end{figure*}

@@

Another interesting result is that we have been able to obtain many
more different users. At the time of this writing we obtained 75
different IPs contributing CPU time to the experiment, giving an
% ¿se podría hacer en el futuro algún tipo de estudio acerca de la procedencia?
% No estoy seguro de que esa relevante, pero bueno...   [Pedro]
% ¿Por país? Por referrer igual sí... También añádelo como issue - JJ
average of around 4 per experiment. However, there is a big variation
in the number of CPUs. There is also a much bigger range in the
performance of these clients. In order to measure that we have
computed the time between two requests from the same IP and plotted it
as an histogram in Figure \ref{fig:intervals}. In that histogram we
can see that there is an almost uniform distribution of intervals,
from the smallest taking a few milliseconds to the biggest with almost
1 second; the setup of the experiment allows it to be accessible from
a very wide range of different IPs. Even within a single IP the
variation can be relatively high, which is shown in a boxplot in
Figure \ref{fig:intervalIP}, sorted by median time. The time between
requests to the server, that is, the time to process 100 generations
of 128 individuals of an evolutionary algorithm to optimize the Trap
function has also very wide range, but even within the same IP,
inter-quartile differences of several hundred milliseconds are not
uncommon. Median for all IPs is 485 milliseconds and average is quite
close, at 496.8 ms.

>>>>

% Rewrite this - JJ
In order to test the volunteer computing environment, a presentation
describing a low cost volunteer computing environment was
created. This presentation was actually delivered in several
conferences (such as IDC 2014, PPSN 2014 and FOSDEM 2015). These
results have been presented in \cite{arxiv:modeling}. In this paper,
although it uses exactly the same evolutionary algorithm client code,
the {\em presentation} has changed: it uses a single page with several
panels showing the progress of the algorithm and how many users have
been contributing to the experiment. To get users, the URL
(\url{http://nodio-jmerelo.rhcloud.com} and
\url{http://nodeo-i.herokuapp.com}) were presented, with a short
message, in Twitter and other social networks, introducing it as a
``demo for evolutionary algorithms'', sometimes in Spanish, some others in
English.

@@

\section{Results}
\label{sec:res}
%
\begin{figure*}[h!tb]
        \centering
        \includegraphics[width=0.95\linewidth]{img/times-base.png}
        \caption{Histogram of time needed to find the solution in the
          command-line client case, with (Y) or without restart (N)
          when the solution is found.}\label{fig:base}
\end{figure*}
%
In order to establish a baseline, we wrote an evolutionary algorithm
in node using the NodEO framework \cite{DBLP:conf/gecco/GuervosVGES14}
in the command line. Conditions, and, indeed, the code, were the same except for the
population, which was set to 512 individuals, and the termination
condition, which included finishing if a sufficient amount of
generations had passed without finding the solution. We did 50 experiments restarting the server,
which contains the pool of best individuals every 100 generations,
when the client had finished without finding the solution, or not;
this second case is more similar to the asynchronous nature of the
browser-based experiment. In fact, no difference was found (using
Wilcoxon test) and the average time to solution was 34.21 seconds.

The experiments done with the first versions of this framework have
already been published elsewhere \cite{arxiv:modeling}. We proved that
up to 20 nodes could participate in a single run if it was
publicized sufficiently, that is, in a Twitter account with enough number
of followers; every experiment was individually announced so
more or less a certain amount of nodes, more than one, at least, were guaranteed. However, it
did not run correctly in mobile phones, the design of the presentation was
not adaptive and thus missed a good amount of possible clients. In
this case the design, which has used the Kube CSS framework, was fully
adaptive and indeed proved to work with mobile phone browsers for
iPhones and Android phones. 
%
\begin{figure}[h!tb]
  \includegraphics[width=0.95\linewidth]{img/openshift-intervals.png}
  \caption{Histogram of the intervals between two requests by the same
  IP for all experiments in the OpenShift server.}  \label{fig:intervals}
\end{figure}
%
\begin{figure*}[h!tb]
  \includegraphics[width=0.90\linewidth]{img/openshift-intervals-IP.png}
  \caption{Boxplot of the intervals between two requests by the same
  IP, by IP, ordered by descending median.}  \label{fig:intervalIP}
\end{figure*}

@@

Another interesting result is that we have been able to obtain many
more different users. At the time of the submission of the first
version of this paper we had obtained 75
different IPs contributing CPU time to the experiment, giving an
% ¿se podría hacer en el futuro algún tipo de estudio acerca de la procedencia?
% No estoy seguro de que esa relevante, pero bueno...   [Pedro]
% ¿Por país? Por referrer igual sí... También añádelo como issue - JJ
average of around 4 per experiment. However, there is a big variation
in the number of CPUs. There is also a much bigger range in the
performance of these clients. In order to measure that we have
computed the time between two requests from the same IP and plotted it
as an histogram in Figure \ref{fig:intervals}. In that histogram we
can see that there is an almost uniform distribution of intervals,
from the smallest taking a few milliseconds to the biggest with almost
1 second; the setup of the experiment allows it to be accessible from
a very wide range of different IPs. Even within a single IP the
variation can be relatively high, which is shown in a boxplot in
Figure \ref{fig:intervalIP}, sorted by median time. The time between
requests to the server, that is, the time to process 100 generations
of 128 individuals of an evolutionary algorithm to optimize the Trap
function has also very wide range, but even within the same IP,
inter-quartile differences of several hundred milliseconds are not
uncommon. Median for all IPs is 485 milliseconds and average is quite
close, at 496.8 ms.
